---
title: "Project SAND: Soils of Agusan del Norte"
format:
  html:
    page-layout: full
    toc: false
---

## Project Overview

Welcome to **Project SAND**. This page displays spatial field locations collected via **QField** joined with physical and chemical soil properties evaluated in the laboratory and a soil profile sketch generated using aqp package in R.

------------------------------------------------------------------------

## Interactive Soil Map

Click on any sampling marker to inspect the associated laboratory parameters. Use the layer control in the top-right corner to toggle between Satellite Imagery and Standard Map views. **(atm: sample data)**

```{r setup, include=FALSE}
# Load required libraries silently
library(sf)
library(leaflet)
library(dplyr)

# Read the merged master data created by process_data.R
master_file <- "data/Soil_Master_Data.gpkg"

if (file.exists(master_file)) {
  soil_data <- st_read(master_file, quiet = TRUE)
} else {
  soil_data <- NULL
}
```

```{r render_map, echo=FALSE, message=FALSE, warning=FALSE}
library(base64enc)
library(leaflet)

if (is.null(soil_data) || nrow(soil_data) == 0) {
  
  htmltools::div(
    style = "padding: 20px; background-color: #f8d7da; color: #721c24; border-radius: 5px;",
    htmltools::h4("No Sampling Data Found"),
    htmltools::p("Please run process_data.R to import and merge field GPKG files before rendering this page.")
  )

} else {

  # Build interactive popup text with embedded Base64 images
  popup_contents <- apply(soil_data, 1, function(row) {
    sample_id <- trimws(as.character(row["Sample_ID"]))
    
    # Path on disk
    local_file_path <- file.path("images", "profiles", paste0(sample_id, ".png"))
    
    # 1. Header & Primary Identification
    html <- paste0(
      "<div style='max-width: 250px; font-family: sans-serif;'>",
      "<strong style='font-size: 15px; color: #2c3e50;'>Sample ID: ", sample_id, "</strong>",
      "<hr style='margin: 6px 0; border: 0; border-top: 1px solid #ccc;'>"
    )
    
    # Add Location if present in row
    if ("Full_Location" %in% names(row) && !is.na(row["Full_Location"])) {
      html <- paste0(html, "<b>Location: </b>", row["Full_Location"], "<br>")
    } else if ("Barangay" %in% names(row) && !is.na(row["Barangay"])) {
      html <- paste0(html, "<b>Barangay: </b>", row["Barangay"], "<br>")
    }
    
    # 2. DYNAMIC LOOP FOR ALL LAB PARAMETERS
    # Non-lab/system columns to exclude from printing
    skip_cols <- c("Sample_ID", "Barangay", "Full_Location", "geometry", "geom")
    lab_cols  <- setdiff(names(row), skip_cols)
    
    for (col in lab_cols) {
      val <- row[col]
      # Only append parameters that have actual values (skips missing/NA lab values)
      if (!is.na(val) && trimws(as.character(val)) != "") {
        # Replace underscores with spaces for clean labels (e.g., Organic_Matter -> Organic Matter)
        pretty_name <- gsub("_", " ", col)
        html <- paste0(html, "<b>", pretty_name, ": </b>", val, "<br>")
      }
    }
    
    # 3. Base64 Profile Sketch Image Embedding
    if (file.exists(local_file_path)) {
      img_raw <- base64enc::dataURI(file = local_file_path, mime = "image/png")
      
      html <- paste0(
        html,
        "<div style='margin-top: 8px;'>",
          "<b>Soil Profile Sketch:</b><br>",
          "<img src='", img_raw, "' style='width: 100%; height: auto; border-radius: 4px; border: 1px solid #ddd; margin-top: 4px;'>",
        "</div>"
      )
    } else {
      html <- paste0(html, "<br><em style='color: #888; font-size: 11px;'>No profile sketch found (Checked: ", local_file_path, ")</em>")
    }
    
    html <- paste0(html, "</div>")
    return(html)
  })

  # Render Leaflet Map
  leaflet(soil_data, width = "100%", height = "700px") %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
    addProviderTiles(providers$OpenStreetMap, group = "Street Map") %>%
    addCircleMarkers(
      radius = 8,
      color = "#ffffff",
      weight = 2,
      fillColor = "#e67e22",
      fillOpacity = 0.9,
      popup = popup_contents,
      label = ~Sample_ID
    ) %>%
    addLayersControl(
      baseGroups = c("Satellite", "Street Map"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    addScaleBar(position = "bottomleft")
}
```

## Summary Information per Location

```{r render_table, echo=FALSE, message=FALSE, warning=FALSE}
library(sf)
library(DT)
library(dplyr)

if (!is.null(soil_data) && nrow(soil_data) > 0) {
  
  # 1. Determine location grouping column (Barangay or Location)
  group_col <- if ("Barangay" %in% names(soil_data)) "Barangay" else "Full_Location"
  
  # 2. Drop spatial geometry and calculate averages per location dynamically
  location_summary <- soil_data %>%
    st_drop_geometry() %>%
    group_by(.data[[group_col]]) %>%
    summarise(
      `Total Samples` = n(),
      # Calculates mean across ALL numeric lab parameters, rounded to 2 decimal places
      across(
        where(is.numeric), 
        ~ round(mean(.x, na.rm = TRUE), 2)
      ),
      .groups = "drop"
    )
  
  # 3. Clean column names (replaces underscores with spaces)
  colnames(location_summary) <- gsub("_", " ", colnames(location_summary))
  
  # 4. Render interactive DataTables widget
  datatable(
    location_summary,
    filter = "top",                  # Adds search/filter boxes per column
    rownames = FALSE,                # Removes default R row numbers
    extensions = 'Buttons',          # Enables CSV/Excel download
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      scrollX = TRUE,                # Allows horizontal scrolling for many parameters
      dom = 'Bfrtip',
      buttons = list(
        'copy', 
        list(extend = 'csv', filename = 'Soil_Summary_By_Location'),
        list(extend = 'excel', filename = 'Soil_Summary_By_Location')
      )
    )
  )

} else {
  htmltools::div(
    style = "padding: 15px; color: #856404; background-color: #fff3cd; border-radius: 4px;",
    "No summary data available to render."
  )
}
```
