---
title: "Project SAND: Soil of Agusan del Norte"
format:
  html:
    page-layout: full
    toc: false
---

## Project Overview

Welcome to **Project SAND**. This page displays spatial field locations collected via **QField** joined with physical and chemical soil properties evaluated in the laboratory and a soil profile sketch generated using aqp package in R.

---

## Interactive Soil Map

Click on any sampling marker to inspect the associated laboratory parameters. Use the layer control in the top-right corner to toggle between Satellite Imagery and Standard Map views.

```{r setup, include=FALSE}
# Load required libraries silently
library(sf)
library(leaflet)
library(dplyr)

# Read the merged master data created by process_data.R
master_file <- "data/Soil_Master_Data.gpkg"

if (file.exists(master_file)) {
  soil_data <- st_read(master_file, quiet = TRUE)
} else {
  soil_data <- NULL
}
```

```{r render_map, echo=FALSE, message=FALSE, warning=FALSE}
library(base64enc)
library(leaflet)

if (is.null(soil_data) || nrow(soil_data) == 0) {
  
  htmltools::div(
    style = "padding: 20px; background-color: #f8d7da; color: #721c24; border-radius: 5px;",
    htmltools::h4("No Sampling Data Found"),
    htmltools::p("Please run process_data.R to import and merge field GPKG files before rendering this page.")
  )

} else {

  # Build interactive popup text with embedded Base64 images
  popup_contents <- apply(soil_data, 1, function(row) {
    sample_id <- trimws(as.character(row["Sample_ID"]))
    
    # Path on disk
    local_file_path <- file.path("images", "profiles", paste0(sample_id, ".png"))
    
    # Header
    html <- paste0(
      "<div style='max-width: 250px; font-family: sans-serif;'>",
      "<strong style='font-size: 15px; color: #2c3e50;'>Sample ID: ", sample_id, "</strong>",
      "<hr style='margin: 6px 0; border: 0; border-top: 1px solid #ccc;'>"
    )
    
    # Add optional lab fields
    if ("pH" %in% names(row) && !is.na(row["pH"])) {
      html <- paste0(html, "<b>pH: </b>", row["pH"], "<br>")
    }
    if ("Organic_Matter" %in% names(row) && !is.na(row["Organic_Matter"])) {
      html <- paste0(html, "<b>Organic Matter: </b>", row["Organic_Matter"], "%<br>")
    }
    
    # Base64 embedding logic
    if (file.exists(local_file_path)) {
      # Read image file and convert directly to a raw Base64 data string
      img_raw <- base64enc::dataURI(file = local_file_path, mime = "image/png")
      
      html <- paste0(
        html,
        "<div style='margin-top: 8px;'>",
          "<b>Soil Profile Sketch:</b><br>",
          "<img src='", img_raw, "' style='width: 100%; height: auto; border-radius: 4px; border: 1px solid #ddd; margin-top: 4px;'>",
        "</div>"
      )
    } else {
      html <- paste0(html, "<br><em style='color: #888;'>No profile sketch found (Checked: ", local_file_path, ")</em>")
    }
    
    html <- paste0(html, "</div>")
    return(html)
  })

  # Render Leaflet Map
  leaflet(soil_data, width = "100%", height = "700px") %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
    addProviderTiles(providers$OpenStreetMap, group = "Street Map") %>%
    addCircleMarkers(
      radius = 8,
      color = "#ffffff",
      weight = 2,
      fillColor = "#e67e22",
      fillOpacity = 0.9,
      popup = popup_contents,
      label = ~Sample_ID
    ) %>%
    addLayersControl(
      baseGroups = c("Satellite", "Street Map"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    addScaleBar(position = "bottomleft")
}
```